---
title: "Summary Statistics"
author: "Aarav Vishesh Dewangan"
date: "08/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary Statistics of 5 Variables:-

```{r summary, echo=FALSE, warning=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
source("../source/summarize.R")
summary_info
```
The 5 variables that we have chosen to dissect the datasets are the following:-

- Min and Max Land Temperature - I have chosen this variable to show the catastrophic
effect of climate change on the planet earth. This variable helps us to see how 
drastically land temperature has changed over the years. Higher temperatures are worsening many types of disasters, including storms, heat waves, floods, and droughts

The maximum land temperature  recorded was 21.32 Celsius in 2011, and the minimum
 land temperature was -5.407 Celsius in 1857. 

- Land and Ocean Average Temperature -  I have chosen this variable to show the catastrophic
effect of climate change on the planet earth.  The difference between this and variable above it is that it includes the ocean temperature as well. The effect of climate change is not only limited to land but it also affecting oceans which can be really harmful for  marine mammals.

The maximum land and ocean average temperature  recorded was 17.611 Celsius in 2015, and the minimum
average land and ocean temperature was 12.475 Celsius in 1857. 

-  Carbon dioxide levels - I have included this variable because increasing carbon 
dioxide level is one of the primary reasons for climate change. So knowing how carbon 
dioxide level is is varying over years is one of the crucial aspects of understanding 
climate change. Increased carbon dioxide emissions are responsible for about two thirds 
of energy imbalances, which causes the earth's temperature to rise by about 
1-1.5 degrees Celsius every year.

The maximum carbon dioxide level was 407.65 ppm in 2016 and the minimum carbon dioxide level was 313.21 ppm in 1958. 

- Average Number of deaths -  I have chosen this variable to show how climate change is responsible 
for causing natural disasters and how they ultimately impact human lives. Knowing this information can strengthen our motive to control climate change.

The mean number of deaths between 1900 and 2021 due to natural disasters is 2842.866. Natural disasters have primarily occurred due to climate change.

- Average Number of Affected people- I have chosen this variable to highlight these large numbers in front of the readers to show that many people might not die due to climate change
but many people are adversely affected by this, which strengthens our motive to show that climate change is catastrophic for the planet earth.

The average number of affected people between 1900 and 2021 due to natural disasters is 882361.9. Natural disasters have primarily occurred due to climate change.

