---
title: "Summary Statistics"
author: "Aarav Vishesh Dewangan"
date: "08/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary Statistics of 5 Variables:-

```{r summary, echo=FALSE, warning=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
source("../source/summarize.R")
summary_info
```
The 5 variables that we have chosen to dissect the datasets are the following:-

- Land Max Temperature - 

The maximum land temperature  recorded was 21.32 Celsius in 2011, and the minimum
 land temperature was -5.407 Celsius in 1857. 

- Land and Ocean Average Temperature -

The maximum land and ocean average temperature  recorded was 17.611 Celsius in 2015, and the minimum
average land and ocean temperature was 12.475 Celsius in 1857. 

-  Carbon dioxide levels -

The maimum carbon dioxide level was 407.65 ppm in 2016 and the minimum carbon dioxide level was 313.21 ppm in 1958. 

- Average Number of deaths -
 
The mean number of deaths between 1900 and 2021 due to natural disasters is 2842.866. Natural disasters have primarily occurred due to climate change.

- Average Number of Affected people

The average number of affected people between 1900 and 2021 due to natural disasters is 882361.9. Natural disasters have primarily occurred due to climate change.

